---
title: "Basket Analysis"
author: "Gustavo Cabelo"

output: 
  html_document:
    highlight: default
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_depth: 2
---

******
# Introduction
## Goal
******

Conduct a market basket analysis through R to discover any interesting relationships (or associations) between customer’s transactions and the item(s) they’ve purchased. These associations can then be used to drive sales-oriented initiatives such as recommender systems like the ones used by Amazon and other eCommerce sites..

******
## Dataset explanation
******

2 documents were used for this analysis:

******
### CSV with online transactions
******
Dataset with **9835 transactions**

******    
### List of products
******
A PDF file listing all the products available:
**125 products** -
**17 products type**

******
# Preprocessing
******

Loading the libraries

```{r,eval=TRUE,echo=TRUE,, error=FALSE, message=FALSE, warning=FALSE,collapse=TRUE}
require(dplyr)
require(magrittr)
require(arules)
require(arulesViz)
require(ggplot2)
require(gridExtra)
require(graphics)
require(corrplot)
```

Loading the given transaction file 

```{r,eval=TRUE,echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}
transactions <- read.transactions(paste0(getwd(),"/Files/ElectronidexTransactions2017.csv"), format = c("basket","single"), sep = ",", rm.duplicates = F)


```

Loading it as dataframe 

```{r,eval=TRUE,echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}
df <- read.csv(paste0(getwd(),"/Files/ElectronidexTransactions2017.csv"), header = F)


```

Understand the transaction file

```{r,eval=TRUE,echo=TRUE}

summary(transactions)

```

Remove the transaction with 0 items

```{r,eval=TRUE,echo=TRUE}
which(size(transactions) == 0)
transactions <- transactions[-which(size(transactions) == 0),]

```

Double check if Rows were removed

```{r,eval=TRUE,echo=TRUE}

which(size(transactions) == 0) # Now we are sure there is no longer transactions with items <=0

```

******
# Data Analysis
## Visualization
******
### Amount of items per transaction
******

>Histogram
```{r,eval=TRUE,echo=TRUE, fig.align='center'}
  hist(size(transactions), breaks = 50, col = 'grey', xlab = "items x transactions", main = "Histogram")


```

>Box Plot
```{r,eval=TRUE,echo=TRUE, fig.align='center'}
boxplot(size(transactions), outline = FALSE, horizontal=TRUE, main="BoxPlot", xlab="items x transactions")

```

    In the histogram and in the BoxPlot we can clearly see that the standard transactions have in average between 2 to 6 products 

******    
### Best Seller items
******
>TOP 10 best seller products
```{r,eval=TRUE,echo=TRUE, fig.align='center'}
itemFrequencyPlot(transactions, type = c("absolute"), topN = 10, ylab = "item frequency")
```


```{r,eval=TRUE,echo=FALSE}
#Vector with the TOP 10 items
top10 <- c(unlist(attributes(head(sort(itemFrequency(transactions, type = c("relative")), decreasing=T), n=10L)))) 
```

>TOP10 best sellers as standalone in the purchase basket

  The following table shows the purchases made as single items compared to the total times this items has been sold.
  
```{r,eval=TRUE,echo=FALSE, warning=FALSE}
## Total transactions
    bestsellers <- as.data.frame(head(sort(itemFrequency(transactions, type = c("absolute")), decreasing=T), n=10L))
    colnames(bestsellers)[1] <- "total"

## Stand alone transactions
  
    standalone = c() # vector to store values
    # Loop over my_matrix
      for(row in 1:10) {
        for(col in 1:10) {
          tbl2 <- transactions[which(size(transactions)==1),]
          tbl2 <-  crossTable(tbl2,sort=T)
          tbl2 <- tbl2[top10,top10]
          tbl2 <- tbl2[row, col]
          
          standalone <- c(standalone,c(tbl2))
        }
      }

    # function to retrun only values different than 0
      isGoodNumber <- function(X) 
      {
        if (X!=0) return(TRUE) else return(FALSE)
      }
    
    # create column in 'bestsellers' with the number of standalone transactions per item 
      isGoodNumber = Vectorize(isGoodNumber)
      standalone <- standalone[isGoodNumber(standalone)]
      bestsellers$standalone <- standalone

      
## Stand alone transactions (%) 
      standalone_perc <- c() 
      # Loop to create % of standalone transactions
      for(row in 1:10) {
        v <- (bestsellers[row,2]/bestsellers[row,1])*100
        v <-  paste(round(v,digits = 2),"%", sep = "")
        standalone_perc <- c(standalone_perc,c(v))
      }

    # create column in 'bestsellers' with the (%) of standalone transactions per item 
    bestsellers$standalone_perc <- standalone_perc

## Product legend (shortner) 
    # create vector to col and row names in the table
      # names <- c(paste0("i",seq_along(1:10)))
    # create column in 'bestsellers' 
      # bestsellers$legend <- names 

bestsellers <- bestsellers[,c(2,3,1)]            
bestsellers

```    

Was observed that in 9 cases out of the TOP 10 Best Seller items the stand alone sells represent less than 10% of the total sells of these items (Table 1), what means that in more than 90% of these sales there is a clear opportunity for cross-selling them with other items.


>Probability of two products appearing together in the basket

  Overview about how the bestseller items correlate in the basket purchases with a second item. Which of them tend to sell as standalone or together with another item.

  The values where the item crossed with itself represent its ocurrence (%) as standalone items in the purchase basket. 
  
```{r,eval=TRUE,echo=FALSE}
  tbl <- transactions[which(size(transactions)>=1),]
  tbl <- crossTable(tbl, 
                     #measure = "support", #<--- comment this row to see the absolute numbers
                     sort=T)
  tbl <- tbl[top10,top10]

  # Loop to replace the diagonal by the standalone value for each product
  for (i in 1:10) tbl[i,i] <- standalone[i]
  
  # Function to transform matrix from absolute to relative(%) values from the total transactions
  support <- function(x){  
  round((x/length(transactions)*100), digits=2)
  }
  
  tbl <- apply(tbl,2,support)
  
  corrplot(tbl, is.corr = FALSE, method = "number", type = "lower")
rm(list = c("standalone", "col", "row", "tbl2", "standalone_perc", "v"))
```


******
# Association rules
## Create rules 
******
**Rules threshold:**

  * support > 0.02 
  * confidence > 0.2
  * lift > 1.5
  
**Metrics Legend:**
    - LHS: Antecedent
    - RHS: Consequent
    - support: indicates how frequently the item / combination appear in the dataset
    - confidence: indicates the number of times the IF/THEN statement on the data are true.
    - lift: indicates the ratio between the observed divided by the expected number of occurences.

```{r,eval=TRUE,echo=TRUE, fig.align='center', results=FALSE}
  rules = apriori(transactions, 
                  parameter=list(support = 0.02, confidence = 0.2))
  
  subrules = rules[quality(rules)$lift > 1.5]
  
  subrules
```

>Visualize rules
The 50 strongest rules after applied the threshold
```{r,eval=TRUE,echo=TRUE, fig.align='center'}

      plot(subrules, measure=c("support","lift"), shading="confidence")
```

- 41 rules with 2 products
- 9 rules with 3 products
```{r,eval=TRUE,echo=TRUE, fig.align='center'}
      plot(subrules, shading="order", control=list(main ="Two-key plot"))
```
      
>List rules
```{r, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
options(width = 800)
  inspect(head(sort(subrules, 
                  by=c("confidence", "support", "lift")),20));
```

In the TOP 20 rules, all of them have either "iMac" or "HP Laptop" as the LHS product, the TOP 2 best-sellers.
Although they are set as the consequence, in case they are were the Antecedent (RHS), the only difference in the metrics would be the "confidence", however it already gives important insights about the strenght of the rules concerning these items in the basket.


******
## Association between 2 Products
******
The next step was to run the “apriori” function in R in order to identify the main basket composition rules associated with the Best Seller items as the RHS and the following item to be added in the basket.

The following metrics were adopted as threshold:

* Parameters applied
    * Frequency > 2%
    *	Confidence > 20%
    *	Minimum quantity of items = 2
    *	Maximum quantity of items = 2

******
### Reccomended products for baskets that contain as first product
******

    **Metrics Legend**
    - LHS: Antecedent
    - RHS: Consequent
    - support: indicates how frequently the item / combination appear in the dataset
    - confidence: indicates the number of times the IF/THEN statement on the data are true.
    - lift: indicates the ratio between the observed divided by the expected number of occurences.

>1- iMac 
```{r,eval=TRUE,echo=TRUE, warning=FALSE, results= FALSE}
  # Create apriori rule
  # For the following items, the code will be hidden. However, to get its results is just to replace "lhs" syntax by the item analyzed.
    rules_iMac = apriori(transactions, 
                        parameter=list(support = 0.02, confidence = 0.2, minlen = 2, maxlen = 2),
                        appearance = list(default="rhs" , "lhs" = c("iMac")))
    
```
```{r set-option, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
options(width = 800)
rules_iMac
    #review the rules and analyse it to filter the rules down to a manageable subset
    inspect(head(sort(rules_iMac, 
                      by=c("confidence", "lift")),10))
```
```{r,eval=TRUE,echo=FALSE, warning=FALSE, fig.align= 'center'}    
  #Plot association rules    
  plot(rules_iMac, method="graph")#, 
       #control=list(type="items"))
  
```

>2- HPLaptop 
```{r,eval=TRUE,echo=FALSE, warning=FALSE, results= FALSE}
# Create apriori rule "HP Laptop"
rules_HPLaptop = apriori(transactions, 
                     parameter=list(support = 0.02, confidence = 0.2, minlen = 2, maxlen = 2),
                     appearance = list(default="rhs" , "lhs" = c("HP Laptop")))

```
```{r, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
options(width = 800)
rules_HPLaptop
#review the rules and analyse it to filter the rules down to a manageable subset
inspect(head(sort(rules_HPLaptop, 
                  by=c("confidence", "lift")),10))
```
```{r,eval=TRUE,echo=FALSE, warning=FALSE, fig.align= 'center'}    
#Plot association rules    
plot(rules_HPLaptop, method="graph")#, 
#control=list(type="items"))

```

>3- CyberPower 
```{r,eval=TRUE,echo=FALSE, warning=FALSE, results= FALSE}
# Create apriori rule "CYBERPOWER Gamer Desktop"
rules_CyberPower = apriori(transactions, 
                         parameter=list(support = 0.02, confidence = 0.2, minlen = 2, maxlen = 2),
                         appearance = list(default="rhs" , "lhs" = c("CYBERPOWER Gamer Desktop")))

```
```{r, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
options(width = 800)
rules_CyberPower
#review the rules and analyse it to filter the rules down to a manageable subset
inspect(head(sort(rules_CyberPower, 
                  by=c("confidence", "lift")),10))
```
```{r,eval=TRUE,echo=FALSE, warning=FALSE, fig.align= 'center'}    
#Plot association rules    
plot(rules_CyberPower, method="graph")#, 
#control=list(type="items"))

```

>4- AppleEarpods 
```{r,eval=TRUE,echo=FALSE, warning=FALSE, results= FALSE}
# Create apriori rule "Apple Earpods"
rules_AppleEarpods = apriori(transactions, 
                         parameter=list(support = 0.02, confidence = 0.2, minlen = 2, maxlen = 2),
                         appearance = list(default="rhs" , "lhs" = c("Apple Earpods")))

```
```{r, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
options(width = 800)
rules_AppleEarpods
#review the rules and analyse it to filter the rules down to a manageable subset
inspect(head(sort(rules_AppleEarpods, 
                  by=c("confidence", "lift")),10))
```
```{r,eval=TRUE,echo=FALSE, warning=FALSE, fig.align= 'center'}    
#Plot association rules    
plot(rules_AppleEarpods, method="graph")#, 
#control=list(type="items"))

```

>5- MacBookAir 
```{r,eval=TRUE,echo=FALSE, warning=FALSE, results= FALSE}
# Create apriori rule "Apple MacBook Air"
rules_MacBookAir = apriori(transactions, 
                         parameter=list(support = 0.02, confidence = 0.15, minlen = 2, maxlen = 2),
                         appearance = list(default="rhs" , "lhs" = c("Apple MacBook Air")))

```
```{r, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
options(width = 800)
rules_MacBookAir
#review the rules and analyse it to filter the rules down to a manageable subset
inspect(head(sort(rules_MacBookAir, 
                  by=c("confidence", "lift")),10))
```
```{r,eval=TRUE,echo=FALSE, warning=FALSE, fig.align= 'center'}    
#Plot association rules    
plot(rules_MacBookAir, method="graph")#, 
#control=list(type="items"))

```

>6- Lenovo Desktop
```{r,eval=TRUE,echo=FALSE, warning=FALSE, results= FALSE}
# Create apriori rule "Lenovo Desktop Computer"
rules_Lenovo = apriori(transactions, 
                         parameter=list(support = 0.02, confidence = 0.2, minlen = 2, maxlen = 2),
                         appearance = list(default="rhs" , "lhs" = c("Lenovo Desktop Computer")))

```
```{r, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
options(width = 800)
rules_Lenovo
#review the rules and analyse it to filter the rules down to a manageable subset
inspect(head(sort(rules_Lenovo, 
                  by=c("confidence", "lift")),10))
```
```{r,eval=TRUE,echo=FALSE, warning=FALSE, fig.align= 'center'}    
#Plot association rules    
plot(rules_Lenovo, method="graph")#, 
#control=list(type="items"))

```

>7- DellDesktop 
```{r,eval=TRUE,echo=FALSE, warning=FALSE, results= FALSE}
# Create apriori rule "Dell Desktop"
rules_DellDesktop = apriori(transactions, 
                         parameter=list(support = 0.02, confidence = 0.2, minlen = 2, maxlen = 2),
                         appearance = list(default="rhs" , "lhs" = c("Dell Desktop")))

```
```{r, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
options(width = 800)
rules_DellDesktop
#review the rules and analyse it to filter the rules down to a manageable subset
inspect(head(sort(rules_DellDesktop, 
                  by=c("confidence", "lift")),10))
```
```{r,eval=TRUE,echo=FALSE, warning=FALSE, fig.align= 'center'}    
#Plot association rules    
plot(rules_DellDesktop, method="graph")#, 
#control=list(type="items"))

```