---
title: "Basket Analysis"
author: "Gustavo Cabelo"

output: 
  html_document:
    highlight: default
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_depth: 2
---

******
# Introduction
## Goal
******

Conduct a market basket analysis through R to discover any interesting relationships (or associations) between customer’s transactions and the item(s) they’ve purchased. These associations can then be used to drive sales-oriented initiatives such as recommender systems like the ones used by Amazon and other eCommerce sites.

******
## Dataset explanation
******

2 documents were used for this analysis:

******
### CSV with online transactions
******
Dataset with **9835 transactions**

******    
### List of products
******
A PDF file listing all the products available:
**125 products** -
**17 products type**

******
# Preprocessing
******

Loading the libraries

```{r,eval=TRUE,echo=TRUE,, error=FALSE, message=FALSE, warning=FALSE,collapse=TRUE}
require(dplyr)
require(magrittr)
require(arules)
require(arulesViz)
require(ggplot2)
require(gridExtra)
require(graphics)
```

Loading the given transaction file 

```{r,eval=TRUE,echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}
transactions <- read.transactions(paste0(getwd(),"/Files/ElectronidexTransactions2017.csv"), format = c("basket","single"), sep = ",", rm.duplicates = F)


```

Loading it as dataframe 

```{r,eval=TRUE,echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}
df <- read.csv(paste0(getwd(),"/Files/ElectronidexTransactions2017.csv"), header = F)


```

Understand the transaction file

```{r,eval=TRUE,echo=TRUE}

summary(transactions)

```

Remove the transaction with 0 items

```{r,eval=TRUE,echo=TRUE}
which(size(transactions) == 0)
transactions <- transactions[-which(size(transactions) == 0),]

```

Double check if Rows were removed

```{r,eval=TRUE,echo=TRUE}

which(size(transactions) == 0) # Now we are sure there is no longer transactions with items <=0

```

******
# Data Analysis
## Visualization
******

<!-- Create Vector of transactional data -->

<!-- ```{r,eval=TRUE,echo=TRUE} -->
<!-- transview <- c(size(transactions)) #This is needed for visualization purposes -->

<!-- ``` -->


Histogram
```{r,eval=TRUE,echo=TRUE}
  hist(size(transactions), breaks = 50, col = 'grey', xlab = "items x transactions", main = "Histogram")
  
 # image: ![Alt text](/Profissional/Ubiqum/Data Analytics - Understanding customers/Programa 2017.02/Course 2 - Predicting Profitability and Customer Preferences 2017/Task 3 - Multiple Regression in R/Git-repository/basket-analysis_files/figure-html/Histogram.png)
  # In the histogram and boxplot is clearly shown that the average amount of items per transaction goes from 2 to 6.

```

Box Plot
```{r,eval=TRUE,echo=TRUE}
boxplot(size(transactions), outline = FALSE, main="Distribution: items x transactions", ylab="Number of items")

```

    In the previous plots we can clearly see that the transactions have in average between 2 to 6 products 
    
#List items####
itemLabels <- as.data.frame(itemLabels(transactions))